---
title: "Fall 2023 Analysis"
author: "Zach Peagler"
date: "2023-10-28"
output: html_document
---


## Load Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse); require(lmerTest); require(RColorBrewer); require(hrbrthemes); require(viridis); require(showtext); require(scico); require(pwr);
```

## Power Analysis

```{r power, echo=FALSE}

pwr.anova.test(k=12, n=12, f=0.4, sig.level=0.05)


```

## Data

### Porometer Data

Filter porometer data to eliminate readings with a leak% greater than 10 and a stomatal conductance greater than 5.
```{r load_p_data, echo=FALSE}
p_data_file <- file.choose()
p_data <- read.csv(p_data_file) %>%
  mutate(Treatment = case_when(RowID==1~"Control",
                               RowID==2~"Soil",
                               RowID==3~"Foliar",
                               RowID==4~"Soil and Foliar",
                               TRUE ~ NA)) %>%
  filter(leak_pct < 10 &
           gsw < 5)

```

### MultispeQ Data

```{r load_m_data, echo=FALSE}
m_data_file<- file.choose() 
m_data <- read.csv(m_data_file) %>%
  mutate(Treatment = case_when(Row=="A"~"Control",
                               Row=="B"~"Soil",
                               Row=="C"~"Foliar",
                               Row=="D"~"Soil and Foliar",
                               TRUE~NA))
```

### Fruit Data

```{r load_f_data, echo=FALSE}
f_data_file<- file.choose()
f_data <- read.csv(f_data_file) %>%
  mutate(Treatment = case_when(Row=="A"~"Control",
                               Row=="B"~"Soil",
                               Row=="C"~"Foliar",
                               Row=="D"~"Soil and Foliar",
                               TRUE ~ NA)) %>%
 mutate(TreatmentNumeric = case_when(Row=="A"~1,
                               Row=="B"~2,
                               Row=="C"~3,
                               Row=="D"~4,
                               TRUE ~ NA)) %>%
  mutate(Presence = 1)
  
```

### Fruit Sugar Data

Filter out penetrometer readings greater than 4.

```{r load_fs_data, echo=FALSE}
fs_data_file<-file.choose()
fs_data <- read.csv(fs_data_file) %>%
    mutate(Treatment = case_when(Row=="A"~"Control",
                               Row=="B"~"Soil",
                               Row=="C"~"Foliar",
                               Row=="D"~"Soil and Foliar",
                               TRUE ~ NA)) %>%
    mutate(SugarAbsolute = (SugarAvg/100)*Weight
                               ) %>%
  filter(Penetrometer < 4 &
         Cluster < 15
         )
```


## Colors

```{r colors, echo=FALSE}

two_colors=scico(2, palette="turku")
four_colors =scico(4, palette="turku")

```

## Fonts

```{r fonts, echo=FALSE}

font_add_google("Open Sans", family = "open")
font_add_google("Montserrat", family = "mont")
showtext_auto()

```

# Analysis

### Porometer Analysis

Plants were measured regularly with a Li-COR LI-600 porometer/fluorometer. Important variables include stomatal conductance (gsw) and photosystem 2 quantum efficiency (PhiPS2).

```{r p_analysis, echo=FALSE}

# Date correction
p_data$Date <- as.Date(p_data$Date, "%m/%d/%y")


# Averages

p_data_avgs <- p_data %>%
  group_by(Treatment) %>%
  summarise_at(vars(gsw), list(name=mean))


# PhiPS2 by date

ggplot(data = p_data, aes(x= Date, y = PhiPS2, color = Treatment)) +
  geom_point()

# GSW by date

ggplot(data = p_data, aes(x= Date, y = gsw, color = Treatment)) +
  geom_point()

# PhiPS2 by treatment

ggplot(data = p_data, aes(x= Treatment, y = PhiPS2, color=Treatment)) +
  geom_boxplot()+
  geom_jitter()

# GSW by treatment

ggplot(data = p_data, aes(x= Treatment, y = gsw, fill=Treatment)) +
  geom_boxplot(alpha=1)+
  #geom_jitter( width=.2, height=0)+
  #geom_violin(alpha=0.5, width=1)+
  scale_fill_manual(values=four_colors)+
  labs(
    title=str_wrap("Figure 1. Stomatal Conductance Across Inoculation Treatments in Salt Stressed Tomato", 40)
  ) +
  ylab(str_wrap("Stomatal Conductance (mol m-2 s-1)", 20))+
  annotate("text", x=3, y=0.4, label="*", size=6)+
  annotate("text", x=2, y=0.225, label="***", size=6)+
  annotate("text", x=4, y=.075, label="***", size=6)+
  theme_minimal()+
  theme(
    legend.position="none",
    text = element_text(size=30, family="mont"),
    axis.title = element_text(size=36, family = "mont", face= "bold"),
    title = element_text(size=30, family="open", face="bold")
  )

# Density figure for stomatal conductance

ggplot(data=p_data, aes(x=gsw, group=Treatment, fill=Treatment))+
  geom_density(adjust=1, alpha=.5)+
  scale_fill_manual(values=four_colors)+
  ylab("Density")+
  xlab("Stomatal Conductance (mol m-2 s-1)")+
  theme_minimal()+
  labs(
    title="Stomatal Conductance Density for Inoculation Treatments in Salt Stressed Tomato"
  )+
  theme(
    legend.position="none",
    text = element_text(size=24, family="mont"),
    axis.title = element_text(size=24, family = "mont", face= "bold"),
    title = element_text(size=20, family="open", face="bold")
  )+
   annotate("text", x=0.35, y=1.8, label="Soil+Foliar", size=4, family="open")+
   annotate("text", x=0.375, y=1.25, label="Foliar", size=4, family="open")+
   annotate("text", x=0.4, y=0.925, label="Soil", size=4, family="open")+
   annotate("text", x=0.55, y=.7, label="Control", size=4, family="open")

## Individual density figures

ggplot(data=p_data, aes(x=gsw, group=Treatment, fill=Treatment))+
  geom_density(adjust=1.5, alpha=.5)+
  scale_fill_manual(values=four_colors)+
  facet_wrap(~Treatment)+
  xlab("Stomatal Conductance (mol m-2 s-1)")+
    ylab("Density")+
    labs(
    title="Stomatal Conductance Density for Inoculation Treatments in Salt Stressed Tomato"
  )+
  theme_minimal()+
  theme(
    legend.position="none",
    axis.title = element_text(family = "open"),
    title = element_text(family="open")
  )

# model with gsw as response, RH as random, and treatment as predictor

summary(model_gsw <- (lmer(
  gsw ~ Treatment + rh_s + (1|PotID),
  data = p_data)))


# model with PhiPS2 as response, Qamb as random, and treatment as predictor

summary(model_PhiPS2 <- (lmer(
  PhiPS2 ~ Treatment + Qamb + (1|PotID),
  data = p_data)))

```

### MultispeQ Analysis

MultispeQ measurements were taken regularly with PhotosynQ MultispeQs. Important variables from this includes Fv'/Fm' and PhiPS2.

```{r m_analysis, echo=FALSE}

#Phi2 analyses

ggplot(data = m_data, aes(x= Treatment, y = Phi2)) +
  geom_boxplot()

summary(model_m_Phi2 <- (lmer(
  Phi2 ~ Treatment + Light.Intensity..PAR. + (1|Pot.ID),
  data = m_data)))

#Fv'/Fm' analyses

ggplot(data = m_data, aes(x= Treatment, y = FvP_over_FmP, fill=Treatment)) +
  geom_violin(alpha=.6)+
  geom_boxplot(alpha=.3, outlier.alpha = 0, width=0.1)+
    scale_fill_manual(values=four_colors)+
    theme_minimal()+
  labs(
    title="FvP/FmP Across Inoculation Treatments in Salt Stressed Tomato"
  )+
    theme(
    legend.position="none",
    axis.title = element_text(family = "open"),
    title = element_text(family="open")
  )


summary(model_m_FvPFmP <- (lmer(
  FvP_over_FmP ~ Treatment + Light.Intensity..PAR. + (1|Pot.ID),
  data = m_data)))


# Leaf thickness

ggplot(data = m_data, aes(x= Treatment, y = leaf_thickness)) +
  geom_boxplot()

summary(model_m_leaf_thickness <- (lmer(
  leaf_thickness ~ Treatment + (1|Pot.ID),
  data = m_data)))

# SPAD

ggplot(data = m_data, aes(x= Treatment, y = SPAD)) +
  geom_boxplot()

summary(model_m_SPAD <- (lmer(
  SPAD ~ Treatment + (1|Pot.ID),
  data = m_data)))



```

### Fruit Analysis

Fruit was collected from tomato plants upon ripening (exit of the breaker stage) and cataloged and weighed. Fruit without blossom-end rot was then brought back to the lab and further analyzed for sugar content.
```{r f_analysis, echo=FALSE}

#Date correction
f_data$Date <- as.Date(f_data$Date, "%m/%d/%y")

# Averages

f_data_avgs <- f_data %>%
  group_by(Row, BER) %>%
  summarise_at(vars(Weight), list(name=mean))

f_data_ber <- f_data %>%
  group_by(Row) %>%
  summarise_at(vars(BER, Presence), list(name=sum))

# F_data sub dataframe creation

f_data_summary <- f_data %>%
#  filter(BER == 0) %>%
  group_by(Treatment, Pot) %>%
  summarize(Final_Weight = sum(Weight))

f_data_control <- f_data %>%
  filter(Treatment == "Control") %>%
  group_by(Treatment, Pot, BER) %>%
  summarize(Final_Weight = sum(Weight))

f_data_s <- f_data %>%
  filter(Treatment == "Soil") %>%
  group_by(Treatment, Pot, BER) %>%
  summarize(Final_Weight = sum(Weight))

f_data_f <- f_data %>%
  filter(Treatment == "Foliar") %>%
  group_by(Treatment, Pot, BER) %>%
  summarize(Final_Weight = sum(Weight))

f_data_sf <- f_data %>%
  filter(Treatment == "Soil and Foliar") %>%
  group_by(Treatment, Pot, BER) %>%
  summarize(Final_Weight = sum(Weight))

# F_data final weight t tests

t.test(Final_Weight ~ BER, data=f_data_control)
t.test(Final_Weight ~ BER, data=f_data_s)
t.test(Final_Weight ~ BER, data=f_data_f)
t.test(Final_Weight ~ BER, data=f_data_sf)

# Final weight by BER across Treatments

ggplot(data = f_data_summary, aes(y = Final_Weight, x = Treatment, fill = as.factor(BER))) +
  geom_boxplot(alpha=1)+
    scale_fill_manual(values=two_colors, labels=c('Undiseased', 'Diseased'))+
  guides(fill= guide_legend(title="Blossom-End Rot"),color="none")+
    theme_minimal()+
  theme(
    legend.position ="top",
    legend.box.background = element_rect(color="gray", linewidth=1),
    text = element_text(size=30, family="mont"),
    axis.title = element_text(size=36, family = "mont", face= "bold"),
    title = element_text(size=30, family="open", face="bold")
  )+
  labs(
    title=str_wrap("Figure 3. Total Fruit Weight Across Inoculation Treatments by Blossom-End Rot Presence", 40)
  )+
  ylab("Total Fruit Weight (g)")+
  annotate("text", x=4.19, y=2750, label="***", size=6)

#BER model

summary(model_f_BER <- (lmer(
  BER ~ Treatment + (1|Pot),
  data = f_data)))

#Fruit analyses

### Fruit weight per treatment violin+boxes
ggplot(data = f_data, aes(x= Treatment, y = Weight, fill=Treatment)) +
  geom_violin(alpha=1)+
  geom_boxplot(alpha=.2, width=0.2, outlier.alpha = 0)+
  labs(
    title=str_wrap("Figure 2. Fruit Weight Across Inoculation Treatments in Salt Stressed Tomato", 40)
  )+
  ylab("Weight(g)")+
  scale_fill_manual(values=four_colors)+
  theme_minimal()+
  theme(
    legend.position = "none",
    text = element_text(size=30, family="mont"),
    axis.title = element_text(size=36, family = "mont", face= "bold"),
    title = element_text(size=30, family="open", face="bold")
  )+
  annotate("text", x=4, y=55, label="***", size=6)+
  annotate("text", x=2, y=55, label="***", size=6)

### Fruit weight by date, jitter, BER color
ggplot(data = f_data, aes(x= Date, y = Weight, color=as.factor(BER))) +
  geom_jitter(width=5, alpha=.9)+
  scale_color_manual(values=two_colors, labels=c('Undiseased', 'Diseased'))+
  guides(color= guide_legend(title="Blossom-End Rot"))+
  labs(
    title="Weight of tomatoes with and without blossom-end rot by date of harvest",
  )+
   ylab("Weight(g)")+
  theme_minimal()+
  theme(
    legend.position=c(.8,.8)
  )

### Fruit weight by date, jitter, Treatment color
ggplot(data = f_data, aes(x= Date, y = Weight, color=Treatment)) +
  geom_jitter(width=5, alpha=.9)+
  scale_color_manual(values=four_colors)+
  labs(
    title="Weight of salt stressed tomatoes across inoculation treatments by date of harvest",
  )+
   ylab("Weight(g)")+
  theme_minimal()+
  theme(
    legend.position=c(.8,.8)
  )

### Density of weight per treatment, together
ggplot(data=f_data, aes(x=Weight, group=Treatment, fill=Treatment))+
  geom_density(adjust=2, alpha=.5)+
  scale_fill_manual(values=four_colors)+
  theme_minimal()+
  theme(
    legend.position="none",
        axis.title = element_text(family = "open"),
    title = element_text(family="open")
  )

### Density of weight per treatment, individual
ggplot(data=f_data, aes(x=Weight, group=Treatment, fill=Treatment))+
  geom_density(adjust=2, alpha=.5)+
  scale_fill_manual(values=four_colors)+
  facet_wrap(~Treatment)+
  theme_minimal()+
  theme(
    legend.position="none",
        axis.title = element_text(family = "open"),
    title = element_text(family="open")
  )

## Models
summary(model_f <- (lmer(
  Weight ~ Treatment + BER + (1|Pot),
  data = f_data)))

summary(model_f <- (lmer(
  Weight ~ Treatment + (1|Pot),
  data = f_data)))

summary(model_f <- (lmer(
  Weight ~ Treatment + Cluster + (1|Pot),
  data = f_data)))

summary(model_f_summary <- (lmer(
  Final_Weight ~ Treatment + (1|Pot),
  data = f_data_summary)))


```



```{r fs_analysis, echo=FALSE}

fs_data$Date <- as.Date(fs_data$Date, "%m/%d/%y")

fs_data_avgs <- fs_data %>%
  group_by(Treatment) %>%
  summarise_at(vars(SugarAvg, Penetrometer), list(name=mean))


fs_data_summary <- fs_data %>%
#  filter(BER == 0) %>%
  group_by(Treatment, Pot) %>%
  summarize(Final_Weight = sum(Weight))

fs_data_summary_sugar <- fs_data %>%
  group_by(Treatment) %>%
  summarize(TotalSugar = sum(SugarAbsolute))

#Fruit sugar analyses

ggplot(data = fs_data, aes(x= Treatment, y = SugarAvg, fill = Treatment)) +
  geom_violin(alpha=1)+
  geom_boxplot(alpha=.2, outlier.alpha=0, width=.25)+
  #geom_jitter(alpha=0.2, width=0.1)+
  labs(
    title=str_wrap("Figure 4. Sugar Concentration Across Inoculation Treatments in Salt Stressed Tomato", 40)
  ) +
  ylab("Sugar%")+
  scale_fill_manual(values=four_colors)+
  theme_minimal()+
    theme(
    legend.position="none",
    text = element_text(size=30, family="mont"),
    axis.title = element_text(size=36, family = "mont", face= "bold"),
    title = element_text(size=30, family="open", face="bold")
  )+
  annotate("text", x=2, y=8.5, label="***", size=6, family="open")+
  annotate("text", x=4, y=9, label="***", size=6, family="open")

ggplot(data = fs_data, aes(x= Treatment, y = SugarAbsolute, fill=Treatment)) +
 geom_violin(alpha=1)+
  geom_boxplot(alpha=.2, outlier.alpha=0, width=.25)+
  labs(
    title=str_wrap("Figure 4. Absolute Sugar Content Across Inoculation Treatments in Salt Stressed Tomato", 40)
  ) +
    ylab("Sugar (g) per Fruit")+
  scale_fill_manual(values=four_colors)+
  theme_minimal()+
    theme(
    legend.position="none",
    text = element_text(size=30, family="mont"),
    axis.title = element_text(size=36, family = "mont", face= "bold"),
    title = element_text(size=30, family="open", face="bold")
  )+
    annotate("text", x=2, y=5, label="***", size=6, family="open")+
    annotate("text", x=4, y=4, label="***", size=6, family="open")

ggplot(data = fs_data, aes(x= Penetrometer, y = SugarAvg, color=Treatment)) +
  geom_point()+
  theme_minimal()

ggplot(data = fs_data, aes(x= Weight, y = SugarAvg, color=Treatment)) +
  geom_point()+
  theme_minimal()+
  facet_wrap(~Treatment)

ggplot(data = fs_data, aes(x= Cluster, y = SugarAvg, color=Treatment)) +
  geom_jitter()+
  theme_minimal()

ggplot(data = fs_data, aes(x=Date, y = SugarAvg, color=Treatment)) +
  geom_jitter(width=10)+
  theme_minimal()

summary(model_fs <- (lmer(
  SugarAvg ~ Treatment + (1|Pot),
  data = fs_data)))

summary(model_fs <- (lmer(
  SugarAbsolute ~ Treatment + (1|Pot),
  data = fs_data)))

summary(model_fs_date <- (lmer(
  SugarAvg ~ Date + (1|Treatment) + (1|Pot),
  data = fs_data)))

summary(model_fs_weight <- (lmer(
  Final_Weight ~ Treatment + (1|Pot),
  data = fs_data_summary)))


## Penetrometer analysis


ggplot(data = fs_data, aes(x=Treatment, y = Penetrometer, fill=Treatment)) +
  geom_violin()+
  geom_boxplot(alpha=0.2, width=0.3, outlier.alpha = 0)+
    labs(
    title=str_wrap("Firmness across inoculation treatments in salt stressed tomato", 40)
  ) +
  ylab("Firmness (kg)")+
  scale_fill_manual(values=four_colors)+
  theme_minimal()+
    theme(
    legend.position="none",
    text = element_text(size=30, family="mont"),
    axis.title = element_text(size=36, family = "mont", face= "bold"),
    title = element_text(size=30, family="open", face="bold")
  )+
  annotate("text", x=4, y=1.825, label="***", size=6, family="open")

summary(model_fs_penetrom <- (lmer(
  Penetrometer ~ Treatment + (1|Pot),
  data = fs_data)))

summary(model_fs_penetrom <- (lmer(
  Penetrometer ~ Treatment + Weight + (1|Pot),
  data = fs_data)))

```
